---
title: "Iris Network analysis"
params: 
  edgelist_path: NULL
  authors_path: NULL
  functions_path: NULL
  figure_output_path: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggraph)
requireNamespace("igraph")
requireNamespace("oaqc")
requireNamespace("concaveman")
source(params$functions_path)
```

This is the code for the network analysis of the Turin author network
For this to work, you must set the working directory to be in the root
of the repository (i.e. `./turin-author-network/`). The calls to `source()`
and the data loading are relative to that position.

```{r, include=FALSE}
# TODO: Some deps for igraph (especially the layout functions) might be
# missing.
# NOTE: To install `concaveman`, to plot the communities, use:
#Sys.setenv(DOWNLOAD_STATIC_LIBV8 = 1)
#install.packages("concaveman")
#library(concaveman)
# this only works if you are using linux.
```

# Loading data
Data comes from the Python pipeline.

```{r}
# Data loading
edgelist <- read_csv(params$edgelist_path)
authors <- read_csv(params$authors_path, na = c("Null", "null", "na", "NA", "None", "none"))
graph <- graph_from_edgelist(edgelist, authors)

fig_prefix <- str_split_1(params$edgelist_path, "/")
fig_prefix <- fig_prefix[length(fig_prefix)]

tee <- function(plot_item, plot_name, width = 12, height = 8, res = 450) {
  png(
    file.path(params$figure_output_path, paste0(fig_prefix, "_", plot_name, ".png")),
    width = width,
    height = height,
    units = "in",
    res = res
  )
  print(plot_item)
  dev.off()
  
  plot_item
}
```

Here we select the departments included in our analysis.

```{r}
## Inspect what departments are available:
#igraph::vertex_attr(graph, "department") |> as.factor() |> levels() |> paste(collapse = "\n") |> cat()

# Select what departments should be included, and what color they should have.
# Optionally (NA ignores), select a new name (e.g. abbreviation) for the
# department, to be shown in the legend.
# The structure is list(old_name = c(new_name, color))
# Note some (like "giurisprudenza") are missing. See above for the full list.
# TODO: Export this to a config file (maybe)?

## A list of "hard science" departments, to use a derogatory term.
# I leave some commented if we want to include them later.
selected_departments <- list(
  "biotecnologie molecolari e scienze per la salute" = c("biotech. mol.", "darkgreen"),
  # NOTE: This department seems like an error - there's just one person in it
  #"centro interdipartimentale di ricerca per le biotecnologie molecolari - mbc" = c("MBC", "limegreen"), 
  "chimica" = c(NA, "orange"),
  #"dental school centro di eccellenza per la ricerca, la didattica e l'assistenza in campo odontostomatologico" = c("dental school", "purple"), 
  "fisica" = c(NA, "blue"),
  "informatica" = c(NA, "lightblue"),
  "matematica giuseppe peano" = c("matematica", "purple"),
  "neuroscienze rita levi montalcini" = c("neuroscienze", "pink"), 
  "oncologia" = c(NA, "brown"),
  "psicologia" = c(NA, "magenta"),
  "scienza e tecnologia del farmaco" = c("farmacologia", "darkorange2"), 
  #"scienze agrarie, forestali e alimentari" = c("agraria", "darkseagreen"),
  #"scienze chirurgiche" = c("chirurgia", "cyan"), 
  "scienze cliniche e biologiche" = c("clinica", "yellow"),
  #"scienze della sanita' pubblica e pediatriche" = c("sanita' pubblica", "cyan4"), 
  #"scienze della terra" = c(NA, "darkseagreen"),
  "scienze della vita e biologia dei sistemi" = c("DBioS", "red"), 
  #"scienze economico-sociali e matematico-statistiche" = c("economia e statistica", "deepskyblue"),
  "scienze mediche" = c("medicina", "darkturquoise")
  #"scienze veterinarie" = c("veterinaria", "darkolivegreen1"),
  #"unknown" = c(NA, "darkgray")
)

# Same as above, but limited to the biological/biomedical areas
selected_departments_biology <- list(
  "biotecnologie molecolari e scienze per la salute" = c("biotech. mol.", "blue"),
  "centro interdipartimentale di ricerca per le biotecnologie molecolari - mbc" = c("MBC", "limegreen"), 
  "neuroscienze rita levi montalcini" = c("neuroscienze", "pink"), 
  "oncologia" = c(NA, "yellow"),
  "psicologia" = c(NA, "magenta"),
  "scienze agrarie, forestali e alimentari" = c("agraria", "purple4"),
  "scienze cliniche e biologiche" = c("clinica", "brown"),
  "scienze della terra" = c(NA, "darkseagreen"),
  "scienze della vita e biologia dei sistemi" = c("DBioS", "red"),
  "scienze veterinarie" = c("veterinaria", "orange")
  #"unknown" = c(NA, "darkgray")
)

# Generate the subgraphs, and add information about communities.
hard_graph <- filter_graph(graph, selected_departments)
hard_graph <- find_communities(hard_graph)

bio_graph <- filter_graph(graph, selected_departments_biology)
bio_graph <- find_communities(bio_graph)

```

## Plotting

First, we make a raw plot of just the network. This is most probably impossible to inspect.

```{r, fig.width=10, fig.height=10}
## PLOTTING GRAPHS
# If 'weights' is NULL the function uses weights in the graph
hard_layout <- graphlayouts::layout_with_sparse_stress(
  hard_graph, weights = NULL, pivots = 25
)
plot_graph(
  hard_graph, selected_departments, hard_layout,
  include_labels = FALSE,
  label_degree_treshold = 40,
  include_communities = FALSE,
  title = "'Hard Science' network graph - Sparse stress layout"
) |> tee("hard_general_graph")

```

```{r}
#faster layout
layout <- graphlayouts::layout_with_sparse_stress(
  bio_graph, weights = NULL, pivots = 25
)
plot_graph(
  bio_graph, selected_departments_biology, layout = layout,
  include_labels = TRUE,
  label_degree_treshold = 40,
  include_communities = FALSE,
  title = "Biological area network - Sparse stress layout"
) |> tee("bio_general_graph")
```


```{r, eval=FALSE, include=FALSE}
# Using igraph::layout_with_dh() is prettier, but takes a very, very long time
layout <- igraph::layout_with_dh(bio_graph)
plot_graph(
  bio_graph, selected_departments_biology, layout = layout,
  include_labels = TRUE,
  label_degree_treshold = 40,
  include_communities = FALSE,
  title = "Biological area network - DH layout"
) |> tee("hard_general_pretty_graph")
```

Now we can inspect single departments:

```{r}
# Inspect single departments
plots <- list(
  "DBios" = inspect_one(
    graph,
    "scienze della vita e biologia dei sistemi",
    "DBios",
    walk = TRUE,
    label_degree_treshold = 0
  ),
  "biotech_mol" = inspect_one(graph, "biotecnologie molecolari e scienze per la salute", "biotech", "green", walk = TRUE, label_degree_treshold = 0),
  "psicologia" = inspect_one(graph, "psicologia", NA, "blue", walk = TRUE),
  "fisica" = inspect_one(graph, "fisica", NA, "orange", label_degree_treshold = 2, walk = TRUE),
  "oncologia" = inspect_one(graph, "oncologia", NA, "purple", label_degree_treshold = 2, walk = TRUE),
  "studi_umanistici" = inspect_one(graph, "studi umanistici", NA, "yellow", label_degree_treshold = 2, walk = TRUE),
  "scienze_mediche" =inspect_one(
    graph, "scienze mediche", NA, "yellow", label_degree_treshold = 2,
    override_communities = 2, walk = TRUE
  )
)

for (i in seq_along(plots)) {
  plot_names <- names(plots)
  tee(plots[i], paste0("individual_plots_", plot_names[i]))
}

```

## Community purity

```{r}
## --- Community purity
bio_purity <- get_community_purity(bio_graph)
plot_purity(bio_purity, selected_departments_biology) |> tee("bio_purity")
plot_purity_dots(bio_purity, selected_departments_biology) |> tee("bio_purity_dots")
plot_purity_dots2(bio_purity, selected_departments_biology) |> tee("bio_purity_dots2")

hard_purity <- get_community_purity(hard_graph)
plot_purity(hard_purity, selected_departments) |> tee("hard_purity")
plot_purity_dots(hard_purity, selected_departments) |> tee("hard_purity_dots")
plot_purity_dots2(hard_purity, selected_departments) |> tee("hard_purity_dots2")
```

## Network Degree distribution

```{r}
## Network Degree
plot_degree_distribution(graph, loglog = TRUE) |> tee("general_degree_distribution")
plot_degree_distribution(hard_graph, loglog = TRUE) |> tee("hard_degree_distribution")
plot_degree_distribution(bio_graph, loglog = TRUE) |> tee("bio_degree_distribution")
```

