---
title: "Iris Network analysis"
params: 
  edgelist_path: NULL
  authors_path: NULL
  functions_path: NULL
  figure_output_path: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggraph)
requireNamespace("igraph")
requireNamespace("oaqc")
requireNamespace("concaveman")

# This source call reads in the `iris_network_analysis_functions.R` contents.
# The path is passed as a param to allow the compilation of the document
# both inside the Docker and locally.
source(params$functions_path)
```

This is the code for the network analysis of the Turin Author Networks.
Please read the README to learn more on what these analysis are doing.

# Loading data
Data comes from the Python pipeline. Please read the general report for more information on the data preprocessing that occurred in Python prior to this analysis.

```{r read-input}
# The edgelist is a csv with source, sink and weight columns.
# `igraph` can just eat it as-is.
edgelist <- read_csv(params$edgelist_path)
authors <- read_csv(
  params$authors_path,
  na = c("Null", "null", "na", "NA", "None", "none")
)
graph <- graph_from_edgelist(edgelist, authors)

fig_prefix <- str_split_1(params$edgelist_path, "/")
fig_prefix <- fig_prefix[length(fig_prefix)]
fig_prefix <- str_split_i(fig_prefix, "_", 2) |> str_split_i("\\.", 1)

# Make figure output dir
fig_dir_path <- file.path(params$figure_output_path, fig_prefix)
dir.create(fig_dir_path, showWarnings = FALSE)

tee <- function(plot_item, plot_name, width = 12, height = 8, res = 450) {
  png(
    file.path(fig_dir_path, paste0(fig_prefix, "_", plot_name, ".png")),
    width = width,
    height = height,
    units = "in",
    res = res
  )
  print(plot_item)
  dev.off()
  
  plot_item
}
```

Here we select the departments included in our analysis.
What were included was an arbitrary choice.
We chose to make two new sub-networks: a "hard-science"s network and a biological area network.
We ignored humanities since they produce weird networks (as they have peculiar publishing processes).

```{r department-notes, include=FALSE, eval=FALSE}
## Some notes for devs on choosing what departments are processed:

# To inspect what departments are available:
#igraph::vertex_attr(graph, "department") |> as.factor() |> levels() |> paste(collapse = "\n") |> cat()
```


```{r department-selection}
# Select what departments should be included, and what color they should have.
# Optionally (NA ignores), select a new name (e.g. abbreviation) for the
# department, to be shown in the legend.
#
# The structure is list(old_name = c(new_name, color))

# TODO: Export this to a config file (maybe)?

## A list of "hard science" departments, to use a derogatory term.
# I leave some commented both to show that they exist, and for ease of access
# if we want to include them later.
selected_departments <- list(
  "biotecnologie molecolari e scienze per la salute" = c("biotech. mol.", "darkgreen"),
  # NOTE: This department seems like an error - there's just one person in it
  #"centro interdipartimentale di ricerca per le biotecnologie molecolari - mbc" = c("MBC", "limegreen"), 
  "chimica" = c(NA, "orange"),
  #"dental school centro di eccellenza per la ricerca, la didattica e l'assistenza in campo odontostomatologico" = c("dental school", "purple"), 
  "fisica" = c(NA, "blue"),
  "informatica" = c(NA, "lightblue"),
  "matematica giuseppe peano" = c("matematica", "purple"),
  "neuroscienze rita levi montalcini" = c("neuroscienze", "pink"), 
  "oncologia" = c(NA, "brown"),
  "psicologia" = c(NA, "magenta"),
  "scienza e tecnologia del farmaco" = c("farmacologia", "darkorange2"), 
  #"scienze agrarie, forestali e alimentari" = c("agraria", "darkseagreen"),
  #"scienze chirurgiche" = c("chirurgia", "cyan"), 
  "scienze cliniche e biologiche" = c("clinica", "yellow"),
  #"scienze della sanita' pubblica e pediatriche" = c("sanita' pubblica", "cyan4"), 
  #"scienze della terra" = c(NA, "darkseagreen"),
  "scienze della vita e biologia dei sistemi" = c("DBioS", "red"), 
  #"scienze economico-sociali e matematico-statistiche" = c("economia e statistica", "deepskyblue"),
  "scienze mediche" = c("medicina", "darkturquoise")
  #"scienze veterinarie" = c("veterinaria", "darkolivegreen1"),
  #"unknown" = c(NA, "darkgray")
)

# Same as above, but limited to the biological/biomedical areas
selected_departments_biology <- list(
  "biotecnologie molecolari e scienze per la salute" = c("biotech. mol.", "blue"),
  "centro interdipartimentale di ricerca per le biotecnologie molecolari - mbc" = c("MBC", "limegreen"), 
  "neuroscienze rita levi montalcini" = c("neuroscienze", "pink"), 
  "oncologia" = c(NA, "yellow"),
  "psicologia" = c(NA, "magenta"),
  "scienze agrarie, forestali e alimentari" = c("agraria", "purple4"),
  "scienze cliniche e biologiche" = c("clinica", "brown"),
  "scienze della terra" = c(NA, "darkseagreen"),
  "scienze della vita e biologia dei sistemi" = c("DBioS", "red"),
  "scienze veterinarie" = c("veterinaria", "orange")
  #"unknown" = c(NA, "darkgray")
)

# Generate the subgraphs, and add information about communities.
hard_graph <- filter_graph(graph, selected_departments)
hard_graph <- find_communities(hard_graph)

bio_graph <- filter_graph(graph, selected_departments_biology)
bio_graph <- find_communities(bio_graph)
```

Note that filtering keeps only the major component of the graph.
This may or may not be desirable, but it helps for visualization and community
detection.

## Network Degree distribution
We begin by describing the network.

```{r network-basic-stats}
get_basic_statistics(graph, "General") |> cat()
# I re-make them here so i can skip filtering the major component.
filter_graph(graph, selected_departments, keep_only_major = FALSE) |>
  get_basic_statistics("Hard-science area") |> cat()
filter_graph(graph, selected_departments_biology, keep_only_major = FALSE) |>
  get_basic_statistics("Bio area") |> cat()
```

Then, we inspect the network degree.
```{r network-degree}
## Network Degree
plot_degree_distribution(graph, loglog = TRUE) |> tee("general_degree_distribution")
plot_degree_distribution(hard_graph, loglog = TRUE) |> tee("hard_degree_distribution")
plot_degree_distribution(bio_graph, loglog = TRUE) |> tee("bio_degree_distribution")
```

## Plotting
First, we make a raw plot of just the network. This is most probably impossible to inspect.

```{r hard-graph-plot}
# If 'weights' is NULL the function uses th weights already in the graph
hard_layout <- graphlayouts::layout_with_sparse_stress(
  hard_graph, weights = NULL, pivots = 25
)
plot_graph(
  hard_graph, selected_departments, hard_layout,
  include_labels = FALSE,
  label_degree_treshold = 40,
  include_communities = FALSE,
  title = "'Hard Science' network graph - Sparse stress layout"
) |> tee("hard_general_graph")

```

```{r bio-graph-plot}
layout <- graphlayouts::layout_with_sparse_stress(
  bio_graph, weights = NULL, pivots = 25
)
plot_graph(
  bio_graph, selected_departments_biology, layout = layout,
  include_labels = TRUE,
  label_degree_treshold = 40,
  include_communities = FALSE,
  title = "Biological area network - Sparse stress layout"
) |> tee("bio_general_graph")
```

```{r pretty-slow-bio-graph, eval=FALSE, include=FALSE}
# Using igraph::layout_with_dh() is prettier, but takes a very, very long time
# This is why this chunk is not included or evaluated, as it slows down
# the analysis by about 10 minutes on my machine.
layout <- igraph::layout_with_dh(bio_graph)
plot_graph(
  bio_graph, selected_departments_biology, layout = layout,
  include_labels = TRUE,
  label_degree_treshold = 40,
  include_communities = FALSE,
  title = "Biological area network - DH layout"
) |> tee("hard_general_pretty_graph")
```

Now we can inspect single departments. The original graph is filtered again to
select just one department, and then plotted.
We use the walker strategy as sometimes - for some reason - the graph is
disconnected and therefore the spinglass approach fails.

```{r single-department-inspection}
# For ease of use, we list all the single departments that we want, and we
# save them later one by one.
# Colors are purely for decoration.
plots <- list(
  "DBios" = inspect_one(
    graph,
    "scienze della vita e biologia dei sistemi",
    "DBios",
    walk = TRUE,
    label_degree_treshold = 0
  ),
  "biotech_mol" = inspect_one(
    graph,
    "biotecnologie molecolari e scienze per la salute",
    "biotech", "green",
    walk = TRUE,
    label_degree_treshold = 0
  ),
  "psicologia" = inspect_one(graph, "psicologia", NA, "blue", walk = TRUE),
  "fisica" = inspect_one(graph, "fisica", NA, "orange", label_degree_treshold = 2, walk = TRUE),
  "oncologia" = inspect_one(graph, "oncologia", NA, "purple", label_degree_treshold = 2, walk = TRUE),
  "scienze_mediche" =inspect_one(
    graph, "scienze mediche", NA, "yellow", label_degree_treshold = 2,
    override_communities = 2, walk = TRUE
  )
)

for (i in seq_along(plots)) {
  plot_names <- names(plots)
  tee(plots[i], paste0("individual_plots_", plot_names[i])) |> print()
}

```

We now look at the MBC and the DBios, which might be interesting.
```{r mbc-bio-graph}
# This will just be used later on for making the dual graph
selected_departments_mbc_bio <- list(
  "biotecnologie molecolari e scienze per la salute" = c("biotech. mol.", "blue"),
  "scienze della vita e biologia dei sistemi" = c("DBioS", "red")
)

mbc_bio_graph <- filter_graph(graph, selected_departments_mbc_bio)
mbc_bio_graph <- find_communities(mbc_bio_graph, number_override = 20)

selected_departments_mbc_bio <- list(
  "biotecnologie molecolari e scienze per la salute" = c("biotech. mol.", "blue"),
  "scienze della vita e biologia dei sistemi" = c("DBioS", "red")
)

layout <- igraph::layout_with_dh(
  mbc_bio_graph
)
plot_graph(
  mbc_bio_graph, selected_departments_mbc_bio, layout = layout,
  include_labels = TRUE,
  label_degree_treshold = 0,
  include_communities = TRUE,
  title = "Biological area network - DH layout"
) |> tee("mbc_bio_graph")
```

## Community purity

We inspect community purity through the purity plots and the purity dot plots.
Please take a look at the descriptions of the generative functions to learn
more on what these plots show.

```{r community-purity}
bio_purity <- get_community_purity(bio_graph)
plot_purity(bio_purity, selected_departments_biology) |> tee("bio_purity")
plot_purity_dots(bio_purity, selected_departments_biology) |> tee("bio_purity_dots")
plot_purity_dots2(bio_purity, selected_departments_biology) |> tee("bio_purity_dots2")

hard_purity <- get_community_purity(hard_graph)
plot_purity(hard_purity, selected_departments) |> tee("hard_purity")
plot_purity_dots(hard_purity, selected_departments) |> tee("hard_purity_dots")
plot_purity_dots2(hard_purity, selected_departments) |> tee("hard_purity_dots2")

# Save the community purities to a file
data_dir <- file.path(params$figure_output_path, fig_prefix, "data")
dir.create(data_dir, showWarnings = FALSE)
write_csv(bio_purity, file = file.path(data_dir, paste0(fig_prefix, "_biological_purity.csv")))
write_csv(hard_purity, file = file.path(data_dir, paste0(fig_prefix, "_hard_purity.csv")))
```

